Amortization of Sub-programs in Simulators

### 2019-09-03 paper/learning/notes.tex
minor

@@

If we have direct access to a rejection sampling loop then it is indeed 
possible to construct the density, up to a constant of proportionality. 
% That is:
% \begin{align*}
% z \sim g(\theta) \\
% do something \\
% \alpha = p_{accept}(z,\theta) \\
% \mathtt{if} Bern(\alpha) = 1: \\
% return z
% \end{align*}
 $x,z \propto f(\theta) p(accept)$.
However, in order to do construct this we need the input $z$, the sample from the proposal $x$ and $p(accept)$ (should use the notation above).
For any rejection sampling block, written in a generic code base, not in the form of R3, this is an incredibly hard 
task to do, because rejection sampling blocks within simulators are not necessarily written in a standard template. 
\textbf{Add examples in a supplement.  Need to demonstrate to YW that this is challenging to do, even if you know where 
the rejection sampling loops are. Hypothesis, RS loops are not written in a templated form}.
y leveraging the PPX protocols previously developed in \cite{baydin2018efficient,gram2019hijacking} we can however, 
extract all variables, other than $P(accept)$ in an automated way.
As without being able to easily extract $p(accept)$ this means that we are unable to 
correctly build the form of the density and thus are limited in our ability to leverage 
powerful inference methods such as MCMC and Variational methods.
Thus, we propose another approach, automating the amortization of rejection sampling loops of type R1, R2 and R3. 
This enables us to learn functions that we can easily sample from and via the amortization process construct the density of the entire
simulator.

@@

\section{Code Examples of Stochastic Processes in Simulators }

The following examples are taken from SHERPA~\cite{gleisberg2009event}
a million-line plus code base used for the modelling of complex physics
simulators at CERN.

>>>>

\subsection{Theory of Nested-programs}

@@

If we have direct access to a rejection sampling loop then it is indeed 
possible to construct the density, up to a constant of proportionality. 
% That is:
% \begin{align*}
% z \sim g(\theta) \\
% do something \\
% \alpha = p_{accept}(z,\theta) \\
% \mathtt{if} Bern(\alpha) = 1: \\
% return z
% \end{align*}
 $x,z \propto f(\theta) p(accept)$.
However, in order to do construct this we need the input $z$, the sample from the proposal $x$ and $p(accept)$ (should use the notation above).
For any rejection sampling block, written in a generic code base, not in the form of R3, this is an incredibly hard 
task to do, because rejection sampling blocks within simulators are not necessarily written in a standard template, see appendix~\ref{app:exampleprograms}.

We leveraging the PPX protocols previously developed in \cite{baydin2018efficient,gram2019hijacking} we can however, 
extract all variables, other than $P(accept)$ in an automated way.
As without being able to easily extract $p(accept)$ this means that we are unable to 
correctly build the form of the density and thus are limited in our ability to leverage 
powerful inference methods such as MCMC and Variational methods.
Thus, we propose another approach, automating the amortization of rejection sampling loops of type R1, R2 and R3. 
This enables us to learn functions that we can easily sample from and via the amortization process construct the density of the entire
simulator.

@@

\section{Code Examples of Stochastic Processes in Simulators }
\label{app:exampleprograms}
The following examples are taken from SHERPA~\cite{gleisberg2009event}
a million-line plus code base used for the modelling of complex physics
simulators at CERN.
